NumPy---Array Computation for Python

### 2020-01-06 body.tex
Some correctness and readability fixes

For many years, NumPy had no dedicated funding, and development time
was mostly contributed freely by students and researchers in their
spare time.  In-person meetings were sponsored off of individual's
research grants, and while industry sponsored contributions were made,
notably those by Mark Wiebe when he was sponsored by Travis Oliphant
at Continuum Analytics, the package's rapid adoption in scientific
computing took place without external investment.

% BIDS -- UCB
% https://www.moore.org/grant-detail?grantId=GBMF5447
% $645,020 in 2016
% https://sloan.org/grant-detail/8222
% $659,359 in 2017
% https://bids.berkeley.edu/news/bids-receives-sloan-foundation-grant-contribute-numpy-development
% http://doi.org/10.5281/zenodo.3585761
% http://doi.org/10.5281/zenodo.3585767
In 2017, NumPy received its first large grants totaling 1.3M USD from the
Gordon \& Betty Moore and the Alfred P. Sloan foundations.
These two grants focus on addressing the technical debt accrued over the years and
on setting in place standards and architecture to encourage more sustainable development.
% CZI -- NumFOCUS/QuanSight
% https://chanzuckerberg.com/eoss/proposals/strengthening-numpys-foundations-growing-beyond-code/
% NumPy and OpenBLAS received $195,000 in 2019
% https://labs.quansight.org/blog/2019/11/numpy-openblas-CZI-grant/
NumPy received a third grant for 195K USD from the Chan Zuckerberg
Initiative at the end of 2019.
This grant focuses on better serving NumPy's large number of beginning
to intermediate level users and on growing the community of NumPy
contributors.
Finally, the project receives around 10K USD annually from Tidelift, which is
typically used to fund documentation writers and website developers.

@@

A vast number of projects are built on NumPy;
these projects are consumers of the NumPy API.
Over the last several years, a growing number of projects are providers of
a \emph{NumPy-like API} and array objects targeting audiences with specialized
needs beyond NumPy's capabilities.
For example, the NumPy API is implemented by several popular tensor computation
libraries including CuPy\footnote{\url{https://cupy.chainer.org/}},
Jax\footnote{\url{https://jax.readthedocs.io/en/latest/jax.numpy.html}},
PyTorch\footnote{\url{https://pytorch.org/tutorials/beginner/blitz/tensor\_tutorial.html}}, and
and Apache MXNet\footnote{\url{https://mxnet.incubator.apache.org/api/python/docs/api/ndarray/index.html}}.
It is also implemented in packages that support sparse arrays
such as \code{scipy.sparse} and \code{pydata.sparse}.
Another notable example is Dask, a library for parallel computing in
Python.  Dask adopts the NumPy API and therefore presents a familiar
interface to existing NumPy users, while adding powerful abilities to
parallelize and distribute tasks.

@@

\section*{Random}

The NumPy random module provides pseudorandom numbers from a wide range of
distributions. In legacy versions of NumPy, simulated random values are produced
by a \code{RandomState} object that: handles seeding and state initialization;
wraps the core pseudorandom number generator based on a 32-bit implementation of
MT19937; interfaces with the underlying code that transforms random bits into
deviates from other distributions; and supplies a singleton instance exposed in
the root of the random module.

@@

Bit generators are simple classes that manage the state of an underlying
pseudorandom number generator. NumPy ships with four bit generators. The default
bit generator is a 64-bit implementation of the Permuted Congruential Generator
\cite{pcg64} (\code{PCG64}). The three other bit generators are a 64-bit version
of the Philox generator\cite{random123} (\code{Philox}), Chris Doty-Humphrey's
Small Fast Chaotic generator\cite{practrand} (\code{SFC64}), and the 32-bit
Mersenne Twister\cite{mt19937} (\code{MT19937}) which has been used in older
versions of NumPy.\footnote{The
\href{https://github.com/bashtage/randomgen}{randomgen project} supplies a wide
range of alternative bit generators such as a cryptographic counter-based
generators (\code{AESCtr}) and generators that expose hardware random number
generators (\code{RDRAND})\cite{randomgen}.} Bit generators export a capsule
containing pointers to functions that produce 64 bits, 32 bits, or a random
double. They also expose CFFI and ctypes interfaces to the same three functions.

>>>>

For many years, NumPy had no dedicated funding, and development time
was mostly contributed freely by students and researchers in their
spare time.  In-person meetings were sponsored off of individual's
research grants, and while industry sponsored contributions were made,
notably those by Mark Wiebe when he was sponsored by Travis Oliphant
at Enthought, the package's rapid adoption in scientific
computing took place without external investment.

% BIDS -- UCB
% https://www.moore.org/grant-detail?grantId=GBMF5447
% $645,020 in 2016
% https://sloan.org/grant-detail/8222
% $659,359 in 2017
% https://bids.berkeley.edu/news/bids-receives-sloan-foundation-grant-contribute-numpy-development
% http://doi.org/10.5281/zenodo.3585761
% http://doi.org/10.5281/zenodo.3585767
In 2017, NumPy received its first large grants totaling 1.3M USD from the
Gordon \& Betty Moore and the Alfred P. Sloan foundations.
These two grants focus on addressing the technical debt accrued over the years and
on setting in place standards and architecture to encourage more sustainable development.
% CZI -- NumFOCUS/QuanSight
% https://chanzuckerberg.com/eoss/proposals/strengthening-numpys-foundations-growing-beyond-code/
% NumPy and OpenBLAS received $195,000 in 2019
% https://labs.quansight.org/blog/2019/11/numpy-openblas-CZI-grant/
NumPy received a third grant for 195K USD from the Chan Zuckerberg
Initiative at the end of 2019.
This grant focuses on better serving NumPy's large number of beginning
to intermediate level users and on growing the community of NumPy
contributors.
Finally, since May 2019 the project receives around 10K USD annually from
Tidelift, which is typically used to fund documentation and website
improvements.

@@

A vast number of projects are built on NumPy;
these projects are consumers of the NumPy API.
Over the last several years, a growing number of projects are providers of
a \emph{NumPy-like API} and array objects targeting audiences with specialized
needs beyond NumPy's capabilities.
For example, the NumPy API is implemented by several popular tensor computation
libraries including CuPy\footnote{\url{https://cupy.chainer.org/}},
JAX\footnote{\url{https://jax.readthedocs.io/en/latest/jax.numpy.html}},
and Apache MXNet\footnote{\url{https://numpy.mxnet.io/}}. 
PyTorch\footnote{\url{https://pytorch.org/tutorials/beginner/blitz/tensor\_tutorial.html}}
and Tensorflow\footnote{\url{https://www.tensorflow.org/tutorials/customization/basics}}
provide tensor APIs with NumPy-inspired semantics.
It is also implemented in packages that support sparse arrays
such as \code{scipy.sparse} and \code{pydata.sparse}.
Another notable example is Dask, a library for parallel computing in
Python.  Dask adopts the NumPy API and therefore presents a familiar
interface to existing NumPy users, while adding powerful abilities to
parallelize and distribute tasks.

@@

\section*{Random number generation}

The NumPy \code{random} module provides pseudorandom numbers from a wide range of
distributions. In legacy versions of NumPy, simulated random values are produced
by a \code{RandomState} object that: handles seeding and state initialization;
wraps the core pseudorandom number generator based on a Mersenne Twister
implementation\footnote{to be precise, the standard 32-bit version of MT19937};
interfaces with the underlying code that transforms random bits into
deviates from other distributions; and supplies a singleton instance exposed in
the root of the random module.

@@

Bit generators are simple classes that manage the state of an underlying
pseudorandom number generator. NumPy ships with four bit generators. The default
bit generator is a 64-bit implementation of the Permuted Congruential Generator
\cite{pcg64} (\code{PCG64}). The three other bit generators are a 64-bit version
of the Philox generator\cite{random123} (\code{Philox}), Chris Doty-Humphrey's
Small Fast Chaotic generator\cite{practrand} (\code{SFC64}), and the 32-bit
Mersenne Twister\cite{mt19937} (\code{MT19937}) which has been used in older
versions of NumPy.\footnote{The
\href{https://github.com/bashtage/randomgen}{randomgen project} supplies a wide
range of alternative bit generators such as a cryptographic counter-based
generators (\code{AESCtr}) and generators that expose hardware random number
generators (\code{RDRAND})\cite{randomgen}.} Bit generators export a capsule
containing pointers to functions that produce
FIXME this doesn't make sense (it's the uniform stream of numbers that matters
rather that "capsule of pointers", also what's "a random double"?):
64 bits, 32 bits, or a random double.
They also expose CFFI and ctypes interfaces to the same three functions.
